I have taken sincere efforts in this project. However, it would not have been possible without the kind support and help of many individuals. I would like to extend my heartfelt thanks to all of them.

I am highly indebted to my Internal Guide \textbf{Prof.\ [Guide Name]} for his/her invaluable guidance and constant supervision, as well as for providing the necessary technical direction regarding the project and for his/her unwavering support in completing this work.

I am grateful to my external guide \textbf{[External Guide Name]} at \textbf{[Company Name]} for providing the industrial perspective, computational resources, and encouragement that were essential for the successful completion of this project.

I would like to express my deepest gratitude to the Head of Department, \textbf{Dr.\ Purvi Prajapati}, and I am also thankful to all my faculty members of Chandubhai S. Patel Institute of Technology for their kind cooperation and encouragement, which helped me in the completion of this project and the preparation of this report.

I also extend my sincere thanks to the open-source community---particularly the developers of Groq, Streamlit, Pillow, and the Llama model family---whose tools and frameworks form the technological backbone of this project.

Last but not the least, I would also like to thank my colleagues, who have cooperated during the preparation of my report and without whom this project would not have been possible. Their ideas, feedback, and constructive criticism helped me significantly improve both the system and this report.

\vspace{0.5cm}
\begin{flushright}
``We may not achieve everything we dream,\\but we cannot achieve anything unless we dream.''
\end{flushright}

\vspace{1.5cm}
\begin{flushright}
\textbf{Viken Hadavani} (ID: XXXXXXXX)
\end{flushright}
